CES analysis for Paul Brockway rebound paper

### 2016-03-18 Articles/Brockway_Rebound_2015/Brockway_2015.Rnw
New CES results for UK 1980 onward.

<<Data>>=
# Create the historicalData data frames
# Set the working directory to the project directory.
historicalData <- read.csv(file = file.path("data", "BrockwayData.csv"))
countries <- c("UK", "US", "CN")
@

@@

<<Models>>=
modelsFileName <- file.path("models", "Brockway2015Models.rds")
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(modelsFileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  models <- lapply(countries, function(coun){
    return(cesModel(iY ~ iK + iL + iU + iYear, filter(historicalData, Country==coun, nest=1:3)))
  })
  names(models) <- countries
  # Save the models.
  saveRDS(models, file = modelsFileName)
}
@

@@

First, look a the UK results.

<<resultsUK, echo=TRUE>>=
summary(models[["UK"]])
naturalCoef(models[["UK"]])
@

@@

@@

<<resultsUS, echo=TRUE>>=
summary(models[["US"]])
naturalCoef(models[["US"]])
@

@@

<<InvestigateUS, echo=TRUE>>=
# Establish some variable names to make code readable
usdata <- filter(historicalData, Country=="US")
usmodel <- models[["US"]]
# Get the boundary model that is near
bmod7 <- (attr(usmodel, "model.attempts"))$`07:*N1-`
# Did Boundary Model 7 converge?
bmod7$converged
# The fitted parameters in bmod7 are as expected,
# i.e., we get NA for both rho and sigma, 
# and we have delta = 1.
naturalCoef(bmod7)
# Calculate the sse for bmod7
sum(resid(bmod7)^2)
# Compare to sse for fitted model
sum(resid(usmodel)^2)
# So, the sse for the winning model is, indeed, 
# less than the sse for the boundary model.
# To investigate why that is happening,
# let's first look at the coefficient for delta
# in the fitted model.
naturalCoef(usmodel)$delta
# Comes back as 1.
# But machine precision is an issue here. 
# Try this:
naturalCoef(usmodel)$delta - 1
# to find that delta is not EXACTLY 1.00000.
# And, this makes a big difference.
# To investigate further, Randy made a function
# that calculates the ln(y) given historical data and 
# some model coefficients.
logces <- function(coefs, data=usdata) {
  coefs <- as.list(coefs)
  with(coefs,
       log(gamma) + 
         lambda * data$iYear + 
         (-1/rho) * (
           log( 
             delta * ( delta_1 * data$iK^(-rho_1) + (1-delta_1) * data$iL^(-rho_1))^(rho/rho_1) +
               (1-delta) * data$iU^(-rho))
         )
  )
}
# Now, calculate sse in many different ways
# First, the boundary model (repeated from above).
sum(resid(bmod7)^2)
# Next from the winning model (repeated from above).
sum(resid(usmodel)^2)
# Using the logces function
sum((log(usdata$iY) - logces(coef(usmodel)))^2)
# Also try with the boundary model.
# First, need to make an array of coefficients that
# matches the full model
bmod7coeffs <- coef(bmod7)
bmod7coeffs[["gamma"]] <- exp(coef(bmod7)[["logscale"]])
bmod7coeffs[["rho"]] <- 2 # value doesn't matter, because delta = 1.
sum((log(usdata$iY) - logces(bmod7coeffs))^2)
# This result matches the sse value from Boundary Model 7.
# The above tests demonstrate that the logces function is correct.
# Let's understand the sensitivity of sse to 
# the fitted coefficients.
# To do so, we'll calculate sse using rounded values 
# of the model coefficients.
# Try rounding to the 16th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 16)))^2)
# The above works as expected. We
# reproduce the sse value from the model itself.
# Try rounding to the 15th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 15)))^2)
# Same sse value as 16th decimal place.
# Again, works as expected.
# Now, try rounding to the 14th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 14)))^2)
# The sse value is slightly different!
# 13th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 13)))^2)
# 12th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 12)))^2)
# Suddenly, we get a big jump in sse value.
# Note that the value of delta above
coef(usmodel)[["delta"]] - 1
# is on the order of 10^-13,
# right at the point we begin to see 
# significant differences in sse.
@

@@

<<resultsCN, echo=TRUE>>=
summary(models[["CN"]])
naturalCoef(models[["CN"]])
@

@@

>>>>

<<Data>>=
# Create the historicalData data frames
# Set the working directory to the project directory.
historicalData <- read.csv(file = file.path("data", "BrockwayData.csv"))
countries <- c("UK", "US", "CN")
# There are places where we need to use data from 1980 onward for the UK only.
UK1980Data <- historicalData %>% filter(Country == "UK") %>% subset(Year >= 1980) %>%
  mutate(
    # Re-normlaize to 1980.
    iYear = iYear - iYear[1],
    iY = iY / iY[1],
    iK = iK / iK[1],
    iL = iL / iL[1],
    iXp = iXp / iXp[1],
    iU = iU / iU[1],
    sF = sF / sF[1],
    sK = sK / sK[1]
  )
@

@@

<<Models>>=
modelsFileName <- file.path("models", "Brockway2015Models.rds")
modelsiXpFileName <- file.path("models", "Brockway2015ModelsiXp.rds")
modelsUK1980FileName <- file.path("models", "Brockway2015ModelsUK1980.rds");
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(modelsFileName)
  modelsiXp <- readRDS(modelsiXpFileName)
  modelsUK1980 <- readRDS(modelsUK1980FileName) # UK using only data from 1980 - 2010.
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  models <- lapply(countries, function(coun){
    return(cesModel(iY ~ iK + iL + iU + iYear, filter(historicalData, Country==coun), nest=1:3))
  })
  names(models) <- countries
  # Save the models.
  saveRDS(models, file = modelsFileName)

  # Now models using iXp instead of iU
  modelsiXp <- lapply(countries, function(coun){
    return(cesModel(iY ~ iK + iL + iXp + iYear, filter(historicalData, Country == coun), nest = 1:3))
  })
  names(modelsiXp) <- countries
  # Save the models.
  saveRDS(modelsiXp, file = modelsiXpFileName)

  # Now run the UK model using only 1980--2010 data
  modelsUK1980 <- list(
    U = cesModel(iY ~ iK + iL + iU + iYear, UK1980Data),
    iXp = cesModel(iY ~ iK + iL + iXp + iYear, UK1980Data)
    )
  saveRDS(modelsUK1980, file = modelsUK1980FileName)
}

# Run these models every time. They are fast.
modelsKL <- lapply(countries, function(coun){
  return(cesModel(iY ~ iK + iL + iYear, filter(historicalData, Country==coun), nest = 1:2))
})
names(modelsKL) <- countries

modelKLUK1980 <- cesModel(iY ~ iK + iL + iYear, UK1980Data, nest = 1:2)
@

@@

First, look a the UK results using all years: 
\Sexpr{historicalData %>% subset(Country == "UK") %>% select(Year) %>% min()}--%
\Sexpr{historicalData %>% subset(Country == "UK") %>% select(Year) %>% max()}.

<<resultsUK, echo=TRUE>>=
summary(models[["UK"]])
naturalCoef(models[["UK"]]) %>% glimpse()
@

@@

Next, UK results using years 
\Sexpr{min(UK1980Data$Year)} -- \Sexpr{max(UK1980Data$Year)}.

<<resultsUK1980, echo=TRUE>>=
summary(modelsUK1980$U)
naturalCoef(modelsUK1980$U) %>% glimpse()
@

The above model with limited years is along a boundary ($\sigma_1 = \infty$) 
and very different from the model using all years.
However, the important thing for rebound is $\rho$,
which is similar between the two models.

@@

<<resultsUS, echo=TRUE>>=
summary(models[["US"]])
naturalCoef(models[["US"]]) %>% glimpse()
@

@@

<<InvestigateUS, echo=TRUE>>=
# Establish some variable names to make code readable
usdata <- filter(historicalData, Country=="US")
usmodel <- models[["US"]]
# Get the boundary model that is near
bmod7 <- (attr(usmodel, "model.attempts"))$`07:*N1-`
# Did Boundary Model 7 converge?
bmod7$converged
# The fitted parameters in bmod7 are as expected,
# i.e., we get NA for both rho and sigma, 
# and we have delta = 1.
naturalCoef(bmod7) %>% glimpse()
# Calculate the sse for bmod7
sum(resid(bmod7)^2)
# Compare to sse for fitted model
sum(resid(usmodel)^2)
# So, the sse for the winning model is, indeed, 
# less than the sse for the boundary model.
# To investigate why that is happening,
# let's first look at the coefficient for delta
# in the fitted model.
naturalCoef(usmodel)$delta
# Comes back as 1.
# But machine precision is an issue here. 
# Try this:
naturalCoef(usmodel)$delta - 1
# to find that delta is not EXACTLY 1.00000.
# And, this makes a big difference.
# To investigate further, Randy made a function
# that calculates the ln(y) given historical data and 
# some model coefficients.
logces <- function(coefs, data=usdata) {
  coefs <- as.list(coefs)
  with(coefs,
       log(gamma) + 
         lambda * data$iYear + 
         (-1/rho) * (
           log( 
             delta * ( delta_1 * data$iK^(-rho_1) + (1-delta_1) * data$iL^(-rho_1))^(rho/rho_1) +
               (1-delta) * data$iU^(-rho))
         )
  )
}
# Now, calculate sse in many different ways
# First, the boundary model (repeated from above).
sum(resid(bmod7)^2)
# Next from the winning model (repeated from above).
sum(resid(usmodel)^2)
# Using the logces function
sum((log(usdata$iY) - logces(coef(usmodel)))^2)
# Also try with the boundary model.
# First, need to make an array of coefficients that
# matches the full model
bmod7coeffs <- coef(bmod7)
bmod7coeffs[["gamma"]] <- exp(coef(bmod7)[["logscale"]])
bmod7coeffs[["rho"]] <- 2 # value doesn't matter, because delta = 1.
sum((log(usdata$iY) - logces(bmod7coeffs))^2)
# This result matches the sse value from Boundary Model 7.
# The above tests demonstrate that the logces function is correct.
# Let's understand the sensitivity of sse to 
# the fitted coefficients.
# To do so, we'll calculate sse using rounded values 
# of the model coefficients.
# Try rounding to the 16th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 16)))^2)
# The above works as expected. We
# reproduce the sse value from the model itself.
# Try rounding to the 15th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 15)))^2)
# Same sse value as 16th decimal place.
# Again, works as expected.
# Now, try rounding to the 14th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 14)))^2)
# The sse value is slightly different!
# 13th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 13)))^2)
# 12th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 12)))^2)
# Suddenly, we get a big jump in sse value.
# Note that the value of delta above
coef(usmodel)[["delta"]] - 1
# is on the order of 10^-13,
# right at the point we begin to see 
# significant differences in sse.
@

@@

<<resultsCN, echo=TRUE>>=
summary(models[["CN"]])
naturalCoef(models[["CN"]]) %>% glimpse()
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Primary Exergy} 
\label{sec:primary_exergy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We can also fit to historical data using quality-adjusted $kl$ and unadjusted $e$
in the form of primary exergy ($iXp$). 
Here are the results.

<<iXp, echo=TRUE>>=
# UK
modelsiXp$UK
# UK starting from 1980
modelsUK1980$iXp
# US
modelsiXp$US
# CN
modelsiXp$CN

# sse for each model

# UK
sum(resid(modelsiXp$UK)^2)
# UK 1980
sum(resid(modelsUK1980$iXp)^2)
# US
sum(resid(modelsiXp$US)^2)
# CN
sum(resid(modelsiXp$CN)^2)

# Compare to sse for models that use iU
# UK
sum(resid(models$UK)^2)
# UK1980
sum(resid(modelsUK1980$U)^2)
# US
sum(resid(models$US)^2)
# CN
sum(resid(models$CN)^2)

@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{$kl$-only models} 
\label{sec:klmodels}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Here are some models fit with only k and l as factors of production.

<<kl, echo=TRUE>>=
# UK
modelsKL$UK
# UK 1980
modelKLUK1980
# US
modelsKL$US
# CN
modelsKL$CN

# sse for each model

# UK
sum(resid(modelsKL$UK)^2)
# UK1980
sum(resid(modelKLUK1980)^2)
# US
sum(resid(modelsKL$US)^2)
# CN
sum(resid(modelsKL$CN)^2)

@
